---
title: "spadjustr: Dual Time-Series Decomposition and Trend Transfer"
author: "Your Name"
output:
  rmarkdown::html_vignette:
    keep_md: true
vignette: >
  %\VignetteIndexEntry{spadjustr: Dual Time-Series Decomposition and Trend Transfer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
library(spadjustr)
library(ggplot2)
library(forecast)
library(gridExtra)
```

# Introduction

This vignette illustrates how to use **spadjustr** to decompose two time series,
swap in the benchmark trend, and visualize the results.

```{r data-prep}
# Example: AirPassengers vs. synthetic benchmark
user_ts <- AirPassengers
set.seed(123)
benchmark_trend <- 100 + 0.5 * seq_along(user_ts)
benchmark_seasonal <- 10 * sin(2 * pi * cycle(user_ts) / 12)
benchmark_noise <- rnorm(length(user_ts), sd = 5)
benchmark_ts <- ts(
  benchmark_trend + benchmark_seasonal + benchmark_noise,
  start = start(user_ts), frequency = frequency(user_ts)
)
```

# Raw Series

```{r raw-series, fig.cap="Raw user and benchmark series"}
p1 <- autoplot(user_ts) + ggtitle("User Series (AirPassengers)") + xlab("Year") + ylab("Passengers")
p2 <- autoplot(benchmark_ts) + ggtitle("Benchmark Series") + xlab("Year") + ylab("Value")
grid.arrange(p1, p2, ncol = 1)
```

# Decomposition and Trend Replacement

We demonstrate several decomposition methods:

```{r methods-plot, fig.cap="Decomposition methods comparison"}
methods <- c("ma12", "stl", "loess")
plots <- lapply(methods, function(m) {
  td <- ts_decomp(user_ts, benchmark_ts, method = m)
  td$plot() + ggtitle(paste("Method:", m))
})
grid.arrange(grobs = plots, ncol = 1)
```

# Detailed STL Example

```{r stl-detail, fig.cap="STL decomposition detail"}
td_stl <- ts_decomp(user_ts, benchmark_ts, method = "stl")
# Combine components and reconstructed series
components <- list(
  Original = user_ts,
  UserTrend = td_stl$decomposed_user$trend,
  BenchTrend = td_stl$decomposed_benchmark$trend,
  Modified = td_stl$modified_user
)
df_comp <- spadjustr:::stack_ts(components)
ggplot(df_comp, aes(x = time, y = value, color = series)) +
  geom_line() + facet_wrap(~ series, scales = "free_y", ncol = 2) +
  ggtitle("STL: Components and Reconstructed Series")
# Prophet decomposition

```{r prophet-example, fig.cap="Prophet method decomposition and trend swap", eval = FALSE}
# The following code demonstrates the prophet method; uncomment to run if 'prophet' is installed:
# td_prop <- ts_decomp(user_ts, benchmark_ts, method = "prophet")
# td_prop$plot() + ggtitle("Method: prophet")
```

# Conclusion

The **spadjustr** package provides an easy R6 interface to decompose time series,
compare trends, and swap trend components. See `?TsDecomp` for more details.
```
